---
title: "p8105_hw1_nel2139"
author: "Nicole Lewis"
date: "2026-09-23"
output: github_document
---

# Section 0: Libraries

This code loads the tidyverse package.
```{r load_packages, message = FALSE}
library(tidyverse)
```

# Problem 1 

## Loading the data

This code loads the `penguins` dataset from the `palmerpenguins` package.

```{r load_data}
data("penguins", package = "palmerpenguins")
```

## Description of the dataset 

The `penguins` dataset contains size measurements for adult foraging penguins observed near Palmer Station, Antarctica, including each penguin's species, island in the Palmer Archipelago, size (flipper length, body mass, and bill dimensions), and sex. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns and includes the following variables: `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`.

The dataset covers three species (`r levels(penguins$species)`) living on three islands (`r levels(penguins$island)`), observed from `r min(penguins$year)` to `r max(penguins$year)`. Bill length ranges from `r min(penguins$bill_length_mm, na.rm = TRUE)` to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm and bill depth ranges from `r min(penguins$bill_depth_mm, na.rm = TRUE)` to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm. Flipper length ranges from `r min(penguins$flipper_length_mm, na.rm = TRUE)` to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm, with a mean value of `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm. Body mass ranges from `r min(penguins$body_mass_g, na.rm = TRUE)` to `r max(penguins$body_mass_g, na.rm = TRUE)` g. Sex is recorded as `r paste(levels(penguins$sex), collapse = " or ")`, and is missing for `r sum(is.na(penguins$sex))` penguins.

## Plots

This code creates a scatterplot of flipper length (y) versus bill length (x), with points colored by species. The second line of code exports the scatterplot to the project directory. 


```{r flip_vs_bill_plot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("flip_vs_bill_plot.png")
```

# Problem 2

The chunk below sets a seed so the same random sample is generated each time to ensure reproducibility and then creates a data frame containing a sample of size 10 from a standard normal distribution constructs the specified logical vector, creates a character vector of length 10, and creates a factor vector of length 10 with three levels. 

```{r problem_2_df}
library(tidyverse)

set.seed(1)

prob2_df = tibble(
  stnorm_samp = rnorm(10, mean = 0, sd = 1),
  stnorm_samp_pos = stnorm_samp > 0, 
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "low", "low", "med", "med", "med", "med", "high", "high", "high"))
)

```


The next chunk of code attempts to take the mean of each variable in the dataframe. 

```{r problem_2_df_means}
mean_stnorm_samp = mean(pull(prob2_df, stnorm_samp))
mean_stnorm_samp_pos = mean(pull(prob2_df, stnorm_samp_pos))
mean_vec_char = mean(pull(prob2_df, vec_char))
mean_vec_factor = mean(pull(prob2_df, vec_factor))


mean_stnorm_samp
mean_stnorm_samp_pos
mean_vec_char
mean_vec_factor
```
Taking the mean works for the numeric and logical variables, `stnorm_samp` and `stnorm_samp_pos`, respectively. The mean of the numeric variable `stnorm_samp` is `r round(mean_stnorm_samp, digits = 2)`, which is the average of the 10 sampled values, this is close to 0, as expected for a standard normal distribution. 
The mean of the logical variable `stnorm_samp_pos` is `r round(mean_stnorm_samp_pos, digits = 2)`, whici is the proportion of sampled values greater than 0. This works because TRUE is treated as 1 and FALSE is treated as 0. 





Taking the mean works for the numeric and logical variables but does not work for the character and factor variables. 
The mean of the numeric variable 






