---
title: "p8105_hw1_nel2139"
author: "Nicole Lewis"
date: "2026-09-23"
output: github_document
---

# Section 0: Libraries

This code loads the tidyverse package.
```{r load_packages, message = FALSE}
library(tidyverse)
```

# Problem 1 

## Loading the data

This code loads the `penguins` dataset from the `palmerpenguins` package.

```{r load_data}
data("penguins", package = "palmerpenguins")
```

## Description of the dataset 

The `penguins` dataset includes measurements for penguin species, island in Palmer Archipelago, size (flipper length, body mass, bill dimensions), and sex. The dataset includes the following variables; `species`, `island`, `bill_len` `bill_dep`, `flipper_len`, `body_mass`, `sex`, and `year`. The species are `r levels(penguins$species)`, and the islands are `r levels(penguins$island)`. Bill length ranges from `r min(penguins$bill_length_mm, na.rm = TRUE)` to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm, bill depth ranges from `r min(penguins$bill_depth_mm, na.rm = TRUE)` to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm, flipper length ranges from `r min(penguins$flipper_length_mm, na.rm = TRUE)` to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm, body mass ranges from `r min(penguins$body_mass_g, na.rm = TRUE)` to `r max(penguins$body_mass_g, na.rm = TRUE)` g. Sex is recorded as `r levels(penguins$sex)`, with some values missing, and the data were collected from `r min(penguins$year)` to `r max(penguins$year)`.

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm.




