---
title: "p8105_hw1_nel2139"
author: "Nicole Lewis"
date: "2026-09-23"
output: github_document
---

# Section 0: Libraries

This code loads the tidyverse package.
```{r load_packages, message = FALSE}
library(tidyverse)
```

# Problem 1 

## Loading the data

This code loads the `penguins` dataset from the `palmerpenguins` package.

```{r load_data}
data("penguins", package = "palmerpenguins")
```

## Description of the dataset 

The `penguins` dataset contains size measurements for adult foraging penguins observed near Palmer Station, Antarctica, including each penguin's species, island in the Palmer Archipelago, size (flipper length, body mass, and bill dimensions), and sex. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns and includes the following variables: `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex`, and `year`.

The dataset covers three species (`r levels(penguins$species)`) living on three islands (`r levels(penguins$island)`), observed from `r min(penguins$year)` to `r max(penguins$year)`. Bill length ranges from `r min(penguins$bill_length_mm, na.rm = TRUE)` to `r max(penguins$bill_length_mm, na.rm = TRUE)` mm and bill depth ranges from `r min(penguins$bill_depth_mm, na.rm = TRUE)` to `r max(penguins$bill_depth_mm, na.rm = TRUE)` mm. Flipper length ranges from `r min(penguins$flipper_length_mm, na.rm = TRUE)` to `r max(penguins$flipper_length_mm, na.rm = TRUE)` mm, with a mean value of `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm. Body mass ranges from `r min(penguins$body_mass_g, na.rm = TRUE)` to `r max(penguins$body_mass_g, na.rm = TRUE)` g. Sex is recorded as `r paste(levels(penguins$sex), collapse = " or ")`, and is missing for `r sum(is.na(penguins$sex))` penguins.

## Plots

This code creates a scatterplot of flipper length (y) versus bill length (x), with points colored by species. The second line of code exports the scatterplot to the project directory. 


```{r flip_vs_bill_plot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("flip_vs_bill_plot.png")
```


# Problem 2

